Recent advances in random matrix theory applied to heavy-tailed multivariate time series

### 2015-09-09 papers/Number1/paper.tex
Revised the paper with Richard David.

@@

\subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et
  al., Soshnikov, Auffiner et al.}
\subsection{$\alpha \in (0,2)$: Our extension of these results; only
  order statistics of $Z_{it}$'s necessary and no sums (proof by the
  technical result)}
To illustrate the validity of the results, we pick the following
distribution function for $Z$ and compare the eigenvalues with
the order statistics of $D_i = \sum_{t=1}^n Z_{it}^2$ and $Z^2$ using
simulated data.
\begin{equation}
  \label{eq:mydist}
  f(x) = {d \over dx}\P(Z \leq x) = \left\{
    \begin{array}{ll}
      {\alpha \over (4|x|)^{\alpha + 1}} & |x| > 1/4 \\
      \alpha & \text{ otherwise }
    \end{array}
  \right.
\end{equation}
This function is shown in figure \ref{fig:densitySim}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{densitySim.pdf}
  \caption{Density function of simulated data. $\alpha=1.6$ is chosen
    for simulation.}
  \label{fig:densitySim}
\end{figure}
Using the simulated data, empirical density functions are obtained for
$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}D_{(i)}$ and
$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}Z^2_{(i)}$. These are shown in
figure \ref{fig:lambda_comparison}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{A total of 20000 pairs are compared in each case.}
  \label{fig:lambda_comparison}
\end{figure}

\subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et
  al., Soshnikov, Auffiner et al.} 
\subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
\subsection{$\alpha \in [2,10/3)$: Our new result for the UN-CENTERED
  sample covariance matrix; only order statistics of $Z_{it}$'s
  necessary and no sums (proof by the technical result)}
\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for
  the UN-CENTERED sample covariance matrix)}

@@

\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample covariance matrix}
\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample covariance matrix}
\subsection{$\alpha \in [2,4)$ and $p/n \to \gamma \in (0,1]$:
  Eigenvalues of the sample covariance matrix using Auffinger's
  result}
\subsection{Point process convergence}

@@

@@

\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample
  autocovariance matrix} 
\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample
  autocovariance matrix}

>>>>

Figure \ref{fig:LamYao} compares the largest eigenvalue of the sum of
time-lagged covariance matrices with the sum of the largest
eigenvalues of individual time-lagged covariance matrices. The data
are rank-transformed return series of 478 stocks selected from the S\&P
500 index.
\begin{figure}[htb!]
  \centering
  \subfigure[] {
    \includegraphics[scale=0.4]{eigen_sum.pdf}
    \label{fig:LamYao:a}
  }
  \subfigure[] {
    \includegraphics[scale=0.4]{eigen_sum_plus.pdf}
    \label{fig:LamYao:a}
  }
  \label{fig:LamYao}
  \caption{The largest eigenvalue of summed time-lagged covariance
    matrices and the sum of the largest eigenvalues of time-lagged
    covariance matrices. In both cases $(C_i)_{lm} =
    (np)^{-2}\sum_{t=1}^{n-i} X_{lt} X_{m,t+i}$}
\end{figure}

@@

\subsection{\texorpdfstring{$\alpha \in (0,2)$}{alpha in (0,2)}:
  Overview of existing results: Davis et al., Soshnikov, Auffiner et
  al.}
\subsection{\texorpdfstring{$\alpha \in (0,2)$}{alpha in (0,2)}: Our
  extension of these results; only order statistics of \texorpdfstring{$Z_{it}$}{Z[it]}'s
  necessary and no sums (proof by the technical result)}
To illustrate the validity of the results, we pick the following
distribution function for $Z$ and compare the eigenvalues with
the order statistics of $D_i = \sum_{t=1}^n Z_{it}^2$ and $Z^2$ using
simulated data.
\begin{equation}
  \label{eq:mydist}
  f(x) = {d \over dx}\P(Z \leq x) = \left\{
    \begin{array}{ll}
      {\alpha \over (4|x|)^{\alpha + 1}} & |x| > 1/4 \\
      \alpha & \text{ otherwise }
    \end{array}
  \right.
\end{equation}
This function is shown in figure \ref{fig:densitySim}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{densitySim.pdf}
  \caption{Density function of simulated data. $\alpha=1.6$ is chosen
    for simulation.}
  \label{fig:densitySim}
\end{figure}
Using the simulated data, empirical density functions are obtained for
$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}D_{(i)}$ and
$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}Z^2_{(i)}$. These are shown in
figure \ref{fig:lambda_comparison}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{A total of 20000 pairs are compared in each case.}
  \label{fig:lambda_comparison}
\end{figure}

\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)}: Overview of existing results: Davis et
  al., Soshnikov, Auffiner et al.}
\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)}: The problem of centering and restrictions on \texorpdfstring{$p_n$}{p[n]}}
\subsection{\texorpdfstring{$\alpha \in [2,10/3)$}{alpha in [2,10/3)}: Our new result for the UN-CENTERED
  sample covariance matrix; only order statistics of \texorpdfstring{\texorpdfstring{$Z_{it}$}{Z[it]}}{Z[it]}'s
  necessary and no sums (proof by the technical result)}
\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: A conjecture with partial proof for
  the UN-CENTERED sample covariance matrix)}

@@

\subsection{\texorpdfstring{$\alpha \in (0,10/3)$}{alpha in (0,10/3)}: Eigenvalues of the sample covariance matrix}
\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: Eigenvalues of the CENTERED sample covariance matrix}
\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)} and
  \texorpdfstring{$p/n \to \gamma \in (0,1]$}{p/n tends to gamma in (0,1]}:
  Eigenvalues of the sample covariance matrix using Auffinger's
  result}
\subsection{Point process convergence}

@@

\subsection{Probability Mass Due To Correlations}
In figure \ref{fig:ProbMass} the distribution function of
${\lambda_{(1)} - \lambda_{(2)} \over \lambda_{(1)}}$ in the cases of
iid data and correlated data, which are described in {\color{red} fill
in here}, is illuatrated. The predicted probability mass at 3/4 is
clearly seen.
\begin{figure}[htb!]
  \centering
  \subfigure[iid data]{
    \includegraphics[scale=0.4]{lambda_1_2_Gaps_iid.pdf}
  }
  \subfigure[Correlated data]{
    \includegraphics[scale=0.4]{lambda_1_2_Gaps.pdf}
  }
  \caption{Distribution function of ${\lambda_{(1)} - \lambda_{(2)} \over
      \lambda_{(1)}}$ in the cases of iid data and correlated
    data. The data matrices have dimension $200 \times 1000$. 1000
    data matrices are simulated in each case. The data have the
    distribution described in \eqref{eq:mydist} with $\alpha=0.6$.}
  \label{fig:ProbMass}
\end{figure}

@@

\subsection{\texorpdfstring{$\alpha \in (0,10/3)$}{alpha in (0,10/3)}: Eigenvalues of the sample
  autocovariance matrix}
\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: Eigenvalues of the CENTERED sample
  autocovariance matrix}
